The Manhattan College community has welcomed me with open arms and provided an environment where I was able to discover my academic interests. For this, I am very grateful for this, and for all the support given to me by my friends. 

My attendance at Manhattan College would not have been possible without the generosity of Mr. and Mrs. O'Malley, to whom I am forever indebted. 

As the first engineer in my family, I relied heavily on advice from all the professors in the Department of Electrical and Computer Engineering at Manhattan College. They have gone beyond providing academic mentorship to nurture my varying interests. In particular, Prof. Nisteruk has witnessed my interests as they changed from signals and systems to linear systems, and finally to quantum computing. I am grateful to him for encouraging me to work outside my comfort levels until I had something to be proud of. His frequent lessons of quantum mechanics outside my regular academic coursework have broadened my perspectives beyond the principles of electrical engineering.

Despite my major, I have always received support from the Department of Mathematics and Computer Science. Several professors in the department took the time to give me independent study courses. I am thankful to Prof.\ Bishop, Prof.\ Boothe, Prof.\ Goldstone, Prof.\ Jura and Prof.\ McCabe for their patience and support. In particular, Prof.\ Bishop and Prof.\ Boothe have nurtured my interest in mathematics and computer science through two rewarding research projects that have helped shape my career significantly.

I also want to acknowledge the support provided to me by the Department of Physics, and particularly by Prof. Liby. He took on the role of advising me willingly and filled the gaps in my knowledge of physical optics. His mentorship has prepared me well for my upcoming graduate studies.

I would like to thank my entire family for their continued support. My academic growth has always been supplemented by their wisdom. In particular, I would like to thank my brother Kidus for constantly reviewing anything I sent his way and my mother Hiwot for the emotional support that she has given me throughout my life at and away from home.

I would also like to thank my girlfriend Mahlet for being a constant source of motivation and encouragement.

Last but certainly not least, I would like to thank God for creating me and allowing me to delve into an exploration of the computing capabilities of quantum phenomena afforded by nature.
